---
title: "SuicideRate_MachineLearning1"
author: "Milica Pajkic, Nuria Piller und Marco Rieder"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Titel

## Suicide Rates in Finland, an analysis based on various factors


```{r, include=FALSE}
library(lubridate)
library(dplyr)
library(readr)
library(data.table)
```


Suicide is a sensible topic and needs to be addressed in an empathical way. 
Behind each number and each case, is a story. 
This paper is analysing data from the statistical office of Finland. Trying to
find evidence what factors are  is one of the goals. The main
data source is this dataset

```{r dataset suicides by monthly data betweend 1971 till 2021}
# data import
d.suicides_monthly <- read_csv("023_12z6_2020m12_20221002-205810.csv")
d.bankruptcy <- read_csv("001_11fb_2022m02_20221004-204725.csv")
d.employment <- read_csv("001_115b_2020_20221004-232300.csv")
d.costOfLviing <- read_csv("001_11xl_2022m08_20221007-084802.csv")
d.businessTrips <- read_csv("111_13q6_2022m04_20221012-204025.csv")
d.averageEarningQ <- read_csv("001_11zu_2022q2_20221012-201102.csv")
d.unemploymentDuration <- read_csv("12ta_2022M08_20221012-204522.csv")

```
Merge data together for analyzing each parameter

```{r}
d.suicide_bankruptcy <- left_join(x=d.suicides_monthly, y= d.bankruptcy, by= "Month")

#remove rows without bankruptcy data (Before 1986)
d.suicide_bankruptcy <- na.omit(d.suicide_bankruptcy)

plot(d.suicide_bankruptcy$`Females Deaths`,d.suicide_bankruptcy$`Bankruptcies instigated, number of employees`)

plot(d.suicide_bankruptcy$`Males Deaths`,d.suicide_bankruptcy$`Bankruptcies instigated, number of employees`)

```

```{r}

d.suicide_costOfLiving <- left_join(x=d.suicides_monthly, y= d.costOfLviing, by= "Month")

#remove rows without bankruptcy data (Before 1986)
d.suicide_costOfLiving <- na.omit(d.suicide_costOfLiving)

plot(x = d.suicide_costOfLiving$`Females Deaths`, y = d.suicide_costOfLiving$`Point figure`)
plot(d.suicide_costOfLiving$`Males Deaths`,d.suicide_costOfLiving$`Point figure`)

```

```{r}
#merge multiple datasets into one major (resolution monthly data)


d.suicide_merged  <-  left_join(d.suicides_monthly, d.bankruptcy, by='Month') %>%
                left_join(., d.costOfLviing, by='Month') %>%
                            left_join(.,d.unemploymentDuration, by='Month') %>%
                            left_join(.,d.businessTrips, by= 'Month')
  
```

