---
title: "Suicide Rates in Finland - Exploration of Correlations"
author: "Milica Pajkic, Nuria Piller und Marco Rieder"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

# 1. Introduction

```{r, include=FALSE}
library(lubridate)
library(dplyr)
library(readr)
library(data.table)
library(tidyverse)
```

Suicide is a complicated topic. And it is still a taboo in modern society. But not only does society struggle with understanding suicidal behavior - also scientists do so.

Studying suicide as a quantitative phenomenon can be especially paradoxical: Suicide is not frequent, and therefore it is hard to model.

> ...in order to prevent suicide more effectively, we must increase the suicide rate. Only then will we be able to reduce the rate!

*David Lester, 1974, quoted in his essay about [Qualitative Research in Suicidology](https://www.researchgate.net/profile/David-Lester-3/publication/308062847_Qualitative_research_in_suicidology_Thoughts_on_Hjelmeland_and_Knizek's_'Why_we_need_qualitative_research_in_suicidology'/links/5c102283a6fdcc494fed8e69/Qualitative-research-in-suicidology-Thoughts-on-Hjelmeland-and-Knizeks-Why-we-need-qualitative-research-in-suicidology.pdf) (2010)*

Still, quantitative methods do what they can with what they have. [They go hand in hand with qualitative methods](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5598439/) to approach the topic of suicide, including a broad range of disciplines from clinical psychology to social sciences to literature. The complexity of examining suicide in a quantitative matter is what made us curious about it.

In this project, we will take a look at the suicide rates of Finland. Why Finland? Scandinavian countries are believed to have higher suicide rates than other European countries. Even though this belief [might not be statistically justified](https://nordics.info/show/artikel/is-suicide-more-common-in-the-nordics), it was enough to made us take a look at the [statistic page of Finland](https://www.stat.fi/index_en.html) and discover, that they have plenty of open source data we can use for our project.

Suicide is a sensible topic and needs to be addressed in an empathetic way. Behind each number and each case, is a story.

```{r load datasets, message=FALSE, warning=FALSE}
# data import
d.suicides_monthly <- read_csv("Data/suicides_monthly.csv") # suicides per month 1971 - 2021
d.bankruptcy <- read_csv("Data/bankruptcy.csv") # bankruptcies per month 1986 - 2020
d.employment <- read_csv("Data/employment.csv") # Yearly Employment statistics 1987 - 2000
d.livingcosts <- read_csv("Data/livingcosts.csv") #living costs per month 1951- Aug 2022
d.trips_abroad <- read_csv("Data/trips_abroad.csv") #trips abroad per month 2012 - April 2022
d.average_earnings <- read_csv("Data/average_earnings.csv") # average earnings by 1/4 year 2000 - 2021
d.unemployment_duration <- read_csv("Data/unemployment_duration.csv") # unemployment duration per month 2006 - Aug 2022

d.consumer_confidence <- read_csv("Data/consumer_Confidence.csv") # Confidence and outlook of people according to their own but also the country, how is their situation and what are they expecting. 


```

Merge data

```{r Merge data and create plots}

#remove columns with redundant information
d.suicides_monthly$`Underlying cause of death (time series classification)` <- NULL

#add total of suicides by adding male and female suicide

d.suicides_monthly$Total_deaths <- d.suicides_monthly$`Males Deaths`+d.suicides_monthly$`Females Deaths`


#merge multiple datasets into one major (resolution monthly data)
#Merging by month for the bankruptcy, costofliving(indexed), duration of 
#unemployment and amount of TripsAbroad


d.s_month <- list(d.suicides_monthly, d.bankruptcy, d.livingcosts, d.unemployment_duration, d.trips_abroad, d.consumer_confidence) %>% reduce(full_join, by= "Month")


 # Plots - erst später mit Kontext - zuerst nur deskriptive Statistiken
# plot(d.suicide_bankruptcy$`Females Deaths`,d.suicide_bankruptcy$`Bankruptcies instigated, number of employees`)

# plot(d.suicide_bankruptcy$`Males Deaths`,d.suicide_bankruptcy$`Bankruptcies instigated, number of employees`)
# 
# plot(x = d.suicide_livingcosts$`Females Deaths`, y = d.suicide_livingcosts$`Point figure`)
# plot(d.suicide_livingcosts$`Males Deaths`,d.suicide_livingcosts$`Point figure`)



```

# 2. Descriptive Statistics

In order to be able to design models which actually make sense, we need some context about the topic itself. Hence, we start with very general descriptive statistics to get a broad overview about the topic. In this chapter, we want to answer to questions:

1.  Did the relative number of suicides increase, considering population increases?

2.  How does suicide relate to Gender and Age?

We work with the following data table indicating the number of suicides per year and by gender 1971 - 2020.

```{r Load data suicides per year and by gender 1971 - 2020}

suicide_1971_2020_gender <- read_delim("Data/suicide_1971-2020_gender.csv", 
     delim = ";", escape_double = FALSE, col_types = cols(Year = col_integer(), 
         Total = col_integer(), Males = col_integer(), 
         Females = col_integer()), trim_ws = TRUE)

```

We will visualize the data in a graph:

```{r Visualization suicides per year and by gender 1971 - 2020}

ggplot(data=suicide_1971_2020_gender, aes(Year)) +
  geom_line(aes(y=Total, colour = "Total")) +
  geom_line(aes(y=Males, colour = "Male")) +
  geom_line(aes(y=Females, colour = "Female"))
 

```

This graph already displays a lot of important information:

-   Around 1990, there was a peak in suicides. Before the peak, the numbers were increasing, and after the peak, they are decreasing.

-   In every point at time, the number of males were significantly higher than the number of females, who committed suicide.

Let us check how the population evolved during this period:

```{r Load and visualize data of poulation growth}

population_and_population_increase_by_sex_1950_2022 <- read_delim("Data/population_and_population_increase_by_sex_1950-2022.csv", 
    delim = ";", escape_double = FALSE, col_types = cols(Year = col_integer(), 
        `Total Population` = col_integer(), 
        `Males Population` = col_integer(), 
        `Females Population` = col_integer(), 
        `Total increase of Population` = col_integer(), 
        `ncrease of Population Males` = col_integer(), 
        `ncrease of Population Females` = col_integer(), 
        `Increase of Population Males` = col_integer(), 
        `Increase of Population Females` = col_integer()), 
    trim_ws = TRUE)

ggplot(data=population_and_population_increase_by_sex_1950_2022, aes(Year))+
  geom_line(aes(y=`Total Population`, colour = "Total")) +
  geom_line(aes(y=`Males Population`, colour = "Male")) +
  geom_line(aes(y=`Females Population`, colour = "Female"))




```

We can see that the population is growing in Finland. We therefore create another graph which shows the number of suicides per 100 000 inhabitants and compare if the trend looks similar.

```{r Create dataframe and visualize suicide rate/100 000 inhabitants}


suicide_rate <- list(population_and_population_increase_by_sex_1950_2022, suicide_1971_2020_gender ) %>% reduce(left_join, by= "Year")


suicide_rate <- suicide_rate %>% 
  mutate(`Cases per 100 000` = suicide_rate$Total / suicide_rate$`Total Population` * 100000)


ggplot(data=suicide_rate, aes(Year))+
   geom_line(aes(y=`Cases per 100 000`))
 
```

This plot shows a similar pattern like the plot with the total population numbers. Hence we assume that the fluctuations of the numbers are not determined by fluctuations of the population, but by other factors.

We now know that the suicide rates in Finland changed a lot since the 70ties - they were rising until the 90ties, and then decreasing. In the following chapters, we will use statistical methods and additional variables to explore eventual correlations.

### 2.2 Other possible explanation for suicide / context of the population at the time

#### Living costs

```{r living costs}

livingcosts <- read_delim("Data/livingcosts.csv", 
     delim = ",", escape_double = FALSE, col_types = cols(Year_Month = col_character(), 
         PointFigure = col_integer()), trim_ws = TRUE)
#I need to split the string Year and Month

livingcosts_right <- livingcosts %>% separate(Month, c("Year", "Month"), sep = "M")
livingcosts_right
typeof(livingcosts_right$`Point figure`)
livingcosts_right$point_figure_n <- as.numeric(livingcosts_right$`Point figure`)
```

```{r livingcosts_right plot}

#livingcosts_right$TIME <- format(as.Date(paste(livingcosts_right$Month, livingcosts_right$Year), format="%m%>"), "%m-%y")

livingcosts_right$TIME <- lubridate::ymd(paste0(livingcosts_right$Year,livingcosts_right$Month,"01")) 

#sc_lc <- plot(livingcosts_right$TIME, livingcosts_right$`Point figure`)

sc_lc1 <- ggplot(data=livingcosts_right, aes(x=TIME, y=livingcosts_right$point_figure_n)) +
  geom_line()
sc_lc1
```

We see that around 1970+ there has been high rise in living costs. We should further consider **unemployment** at the same time.\
[**What we can look at is maybe the percentage of living cost rise to the previous year.**]{.underline}

#### Employment

```{r}
employment <- read_delim("Data/employment.csv", 
     delim = ",")
names(employment)
```

#### Unemployment

```{r}
unemployment <- read_delim("Data/unemployment_duration.csv", 
     delim = ",")

unemployment_right <- unemployment %>% separate(Month, c("Year", "Month"), sep = "M")
unemployment_right

unemployment_right$TIME <- lubridate::ymd(paste0(unemployment_right$Year,unemployment_right$Month,"01")) 
unemployment_right

typeof(unemployment_right$`Unemployed jobseekers on calculation date (number) Total`)

#Rename the Variables
rename(unemployment_right, Total = `Unemployed jobseekers on calculation date (number) Total`)
```

Now let's plot the data on unemployment. Because the data is categorically, we are going to do a bar plot

```{r}
#total
p<-ggplot(data=unemployment_right, aes(x=TIME, y=unemployment_right$`Unemployed jobseekers on calculation date (number) Total`)) +
  geom_bar(stat="identity")
p


#more than 2 years unemployed
p1<-ggplot(data=unemployment_right, aes(x=TIME, y=unemployment_right$`Unemployed jobseekers on calculation date (number) 2-3 years`)) +
  geom_bar(stat="identity")
p1

```

#### Bankruptcy

```{r}
bankruptcy <- read_delim("Data/bankruptcy.csv", 
     delim = ",")

bankruptcy_right <- bankruptcy %>% separate(Month, c("Year", "Month"), sep = "M")
bankruptcy_right

bankruptcy_right$TIME <- lubridate::ymd(paste0(bankruptcy_right$Year,bankruptcy_right$Month,"01")) 
bankruptcy_right
bankruptcy_right %>% glimpse()
```

```{r}
bank_p <- ggplot(data=bankruptcy_right, aes(x=TIME, y=bankruptcy_right$`Bankruptcies instigated, number of employees`)) +
  geom_line()
bank_p

bank_p1<-ggplot(data=bankruptcy_right, aes(x=TIME, y=bankruptcy_right$`Bankruptcies instigated, number of employees`)) +
  geom_bar(stat="identity")
bank_p1
```

We see a high rise between 1990 and 1997

One thing we could look at is the **difference** to the previous year/month

```{r make an difference column}
#bankruptcy_right %>%
  #arrange(TIME, `Bankruptcies instigated, number of employees`) %>%
  #group_by(TIME) %>%
  #mutate(diff = `Bankruptcies instigated, number of employees` - 
        #   lag(`Bankruptcies instigated, number of employees`))
bankruptcy_diff <- diff(bankruptcy_right$`Bankruptcies instigated, number of employees`)
bankruptcy_diff <- c(NA, bankruptcy_diff)
bankruptcy_right$difference <- bankruptcy_diff
```

```{r plot the diff}
bank_diff_p1<-ggplot(data=bankruptcy_right, aes(x=TIME, y=difference)) +
  geom_bar(stat="identity")
bank_diff_p1
summary(bankruptcy_right$difference)
```

```{r}
ggplot(bankruptcy_right, aes(x=TIME, y=difference)) + 
  geom_bar(stat = "identity", width=60)
```

# 3. *Who* is committing suicide? Linear Model with the variables Age and Gender

We first start with a linear regression model. What we want to know: Is there a specific gender and age spectrum, in which the suicide cases appear more frequent? And if so - how do these correlations evolve over time?

To analyse this question, we use a data set indicating the gender of the suicide cases between 1998 and 2020. What is more, the cases are clustered in age groups by 10 years. This is the only data set we found which provides age related information about the cases, which is the reason why we go with that.

```{r Import Dataset Agegroups and Gender 1998 - 2020, message=FALSE, warning=FALSE, include=FALSE}
library(readr)
suicide_yearly_age_gender_1998_2020 <- read_delim("Data/suicide_yearly_age_gender_1998-2020.csv", 
    delim = ";", escape_double = FALSE, col_types = cols(Year = col_date(format = "%Y"), 
        Total = col_integer(), `Okt 14` = col_integer(), 
        `15 - 19` = col_integer(), `20 - 24` = col_integer(), 
        `25 - 29` = col_integer(), `30 - 34` = col_integer(), 
        `35 - 39` = col_integer(), `40 - 44` = col_integer(), 
        `45 - 49` = col_integer(), `50 - 54` = col_integer(), 
        `55 - 59` = col_integer(), `60 - 64` = col_integer(), 
        `65 - 69` = col_integer(), `70 - 74` = col_integer(), 
        `75 - 79` = col_integer(), `80 - 84` = col_integer(), 
        `85 - 89` = col_integer(), `90 - 94` = col_integer(), 
        `95 -` = col_integer(), Males = col_integer(), 
        Females = col_integer()), trim_ws = TRUE)
View(suicide_yearly_age_gender_1998_2020)
```

```{r}

ggplot(data=suicide_yearly_age_gender_1998_2020, aes(Year)) +
  geom_line(aes(y=`Okt 14`, colour = "0 - 14")) +
  geom_line(aes(y=`15 - 19`, colour = "15 - 19")) +
  geom_line(aes(y=`20 - 24`, colour = "20 - 24")) +
  geom_line(aes(y=`25 - 29`, colour = "25 - 29")) +
  geom_line(aes(y=`30 - 34`, colour = "30 - 34"))
# 
# `15 - 19`
# `25 - 29`
# `35 - 39`
# `45 - 49`
# `55 - 59`
# `65 - 69`
# `75 - 79`
# `85 - 89`
```

\

# 

### Content coming soon:

### 

HEAD *-- a Generalised Linear Model with family set to Poisson -- a Generalised Linear Model with family set to Binomial -- a Generalised Additive Model -- a Neural Network -- a Support Vector Machine - solve an optimization problem* ======= - *-- a Generalised Linear Model with family set to Poisson*

-   *-- a Generalised Linear Model with family set to Binomial*

-   *-- a Generalised Additive Model -- a Neural Network*

-   *-- a Support Vector Machine*

## Kapitel Mili

```{r test mili}

```

```{r Poisson Model}

```
