---
title: "SuicideRate_MachineLearning1"
author: "Milica Pajkic, Nuria Piller und Marco Rieder"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Suicide Rates in Finland, an analysis based on various factors


```{r, include=FALSE}
library(lubridate)
library(dplyr)
library(readr)
library(data.table)
```


Suicide is a sensible topic and needs to be addressed in an empathical way. 
Behind each number and each case, is a story. 
This paper is analysing data from the statistical office of Finland. Trying to
find evidence what factors are statisticly related is one of the goals. The main
data source is this dataset

```{r dataset suicides by monthly data betweend 1971 till 2021}

# data import
d.suicides_monthly <- read_csv("023_12z6_2020m12_20221002-205810.csv")
d.bancrupcy <- read_csv("001_11fb_2022m02_20221004-204725.csv")
d.employment <- read_csv("001_115b_2020_20221004-232300.csv")

# Summary statistic of the imported files

summary(d.suicides_monthly)
summary(d.bancrupcy)
summary(d.employment)


```
Merge data together for analyzing each parameter

```{r}
d.suicide_bancrupcy <- left_join(x=d.suicides_monthly, y= d.bancrupcy, by= "Month")

#remove rows without bankruptcy data (Before 1986)
d.suicide_bancrupcy <- na.omit(d.suicide_bancrupcy)

plot(d.suicide_bancrupcy$`Females Deaths`,d.suicide_bancrupcy$`Bankruptcies instigated, number of employees`)

plot(d.suicide_bancrupcy$`Males Deaths`,d.suicide_bancrupcy$`Bankruptcies instigated, number of employees`)

```

