---
title: "European Survey - What is the political sentiment in German speaking Countries Before Covid?"
author: "Milica Pajkic and Marco Rieder"
date: "`r Sys.Date()`"
code_folding: hide
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

# Getting the Data and installing libraries

```{r, include=FALSE}
library(lubridate)
library(multcomp)
library(dplyr)
library(readr)
library(data.table)
library(tidyverse)
theme_set(theme_bw())
library(naniar)
library(neuralnet)
library(nnet)
library(caret)
library(gamlss.add)
library(ggplot2)
library(mgcv)

```

```{r}
ess9 <- read_csv("ESS9.csv")

unique(ess9$name)
unique(ess9$gndr)
unique(ess9$cntry)
```

```{r making subsets of the whole dataset to only german speaking countries}
ess9_de <- subset(ess9, cntry == "DE", select =dweight:domain)
ess9_ch <- subset(ess9, cntry == "CH", select =dweight:domain)
ess9_at <- subset(ess9, cntry == "AT", select =dweight:domain)
```

# 1. Introduction

The European Social Survey (ESS) is a large scale survey conducted in over 38 countries within Europe. Focusing on public attitudes and values and changes within time. This paper is evaluating the ninth version of the survey (ESS9) form 2018.

## 1.1 Selection of parameters

The survey is really comprehensive and consists out of 572 variables. Some of them are related to others and only answered by a subset of participants. 

## 1.2 Goal

# 2. Descriptive Statistics

First 

```{r }
summary(ess9$nwspol)
sum(is.na(ess9$nwspol))
```

```{r echo=FALSE}
summary(ess9_de$gndr)
sum(is.na(ess9_de$gndr))
typeof(ess9_de$nwspol)
```

```{r}
ggplot(ess9_de, aes(nwspol, 1)) +
  geom_bar(stat = "identity") + 
  xlim(c(0, 200))
```

```{r}
boxplot(ess9_de$nwspol, outline=FALSE)
```

### **Was muss man alles für eine deskriptive Analyse machen?**

```{r lineare Regression}
typeof(ess9_de$eduade3)
summary(ess9_de$eduade3)
sum(is.na(ess9_de$eduade3))
```

# 3. Models

## 3.1 Linear Model

How much politics are you watching = DV\
education level by years = IV\
happy with government = IV\
Gender = IV\
Age = IV


```{r linear Regression}
lm_news.1 <- lm(nwspol ~ eduade3, data = ess9_de)
summary(lm_news.1)
```

## 3.2 GLM Poisson
### 3.2.1Data preparation

```{r, data preparation, include=TRUE}
# Setting NA values correct for the used parameters replace with NA replace function from naniar package. As each variable has a different NA setting, it is specified for each variable.

ess9_prep <- ess9 %>%
  replace_with_na(replace = list(nwspol =
                                   c(6666, 7777, 8888, 9999),
                                 polintr = c(7,8,9),
                                 trstprl = c(77,88,99),
                                 eisced = c(77,88,99),
                                 eduyrs = c(77, 88, 99),
                                 stfeco = c(77, 88, 99),
                                 stfgov = c(77, 88, 99),
                                 stfdem = c(77, 88, 99),
                                 gndr = 9,
                                 agea = 999,
                                 rlgdgr = c(77, 88, 99),
                                 netustm = c(6666, 7777, 8888, 9999),
                                 netusoft = c(7,8,9),
                                 psppsgva = c(7,8,9),
                                 grspnum = c(666666666,777777777,888888888,999999999),
                                 infqbst = c(6,7,8,9)
                                 
                                 ))

#testing if change has worked and how many NA we have - for four examples

sum(is.na(ess9_prep$nwspol))
sum(is.na(ess9_prep$eduyrs))
sum(is.na(ess9_prep$stfgov))
sum(is.na(ess9_prep$netusoft))



# The variable "grspnum" is depended on the variable "infqbst." As "infqbst" sets the period of the payment and "grpsnum" is the payment amount for the defined period. To standardize it on a yearly gross payment it needs a nested ifelse statement. If "infqbst" is one, we multiple it by 52 weeks of the year, if it is a two we multiple it by 12, and if it something else we keep it as it is. 

ess9_prep$yrpy <- ifelse(ess9_prep$infqbst %in% 1,
                                ess9_prep$grspnum*52,  
                            ifelse(ess9_prep$infqbst %in% 2,
                                ess9_prep$grspnum*12, 
                                ess9_prep$grspnum))
```


```{r, data preparation, include=TRUE}
# defining subset for study and models. These factors will be used in the Poisson and Neural Network models below. 
ess9_subset <- c("nwspol", "polintr", "trstprl", "eisced", "eduyrs","stfeco", "stfgov", "stfdem", "gndr", "agea","rlgdgr","netusoft", "psppsgva", "yrpy")
```


```{r, data preparation, include=TRUE}
#subsetting the dataset with the defined parameters
ess9_prep <- ess9_prep[ess9_subset]
```


```{r, data preparation, include=TRUE}
# drop NA for the columns used for the poisson model
ess9_prep <- ess9_prep %>% drop_na((ess9_subset))
#keep data for AMM part as numeric values
ess9_prep_num <- ess9_prep
```


```{r, data preparation, include=TRUE}
# set of parameters which are categorical variables
factors <- c("polintr","trstprl", "eisced", "stfeco", "stfgov", "stfdem", "gndr", "rlgdgr", "psppsgva", "netusoft")

# applying factorization of variables
ess9_prep[factors] <- lapply(ess9_prep[factors], as.factor)
#show data after preparation
str(ess9_prep)

```

The media consumption information from the survey are measured in minutes. These Information are count data. In this study a GLM Poisson model is applied. It is used to simulate the different media consumption by Gender. 
```{r}
# Use cleaned and factorized data for the Poisson model
ess9_poisson <- ess9_prep

ggplot(data = ess9_poisson, mapping = aes(y = nwspol,x = gndr)) +
  geom_boxplot() +
  ggtitle("Media consumption in minutes compared to Gender (Zoomed In)")+
  ylab("media consumption in minutes") +
  xlab("Gender") +
  coord_cartesian(ylim = quantile(ess9_poisson$nwspol, c(0.1, 0.9))) # large outliers are cropt out to indicate variance of male and female

```

### 3.2.2 Fitting the poisson model
```{r}
# fitting the model
glmP_news <- glm(nwspol ~ polintr + eisced + trstprl+ eduyrs + netusoft + stfeco + stfgov +
                   stfdem + gndr + agea + rlgdgr +yrpy,
                 family = "poisson", ## we specify the distribution!
                 data = ess9_poisson)

summary(glmP_news)

coef(glmP_news)

exp(coef(glmP_news)["(Intercept)"])

```

### 3.2.3 Simulation
Now we simulate with our model the average media consumption for Females and Males.
```{r poisson}


set.seed(2) # for reproducibility
#simulate the data from glm poisson
sim.data.ess9.Poisson <- simulate(glmP_news)
##

NROW(sim.data.ess9.Poisson)
head(sim.data.ess9.Poisson)

#plot data as bar plot, with gender as factor to get two groups to compare, without factoring only one plot would show.
ggplot(mapping = aes(y = sim.data.ess9.Poisson$sim_1,
                     x = ess9_poisson$gndr)) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  ylab("simulated media consumption \n(assuming Poisson dist)") +
  xlab("Gender") +
  coord_cartesian(ylim = quantile(ess9_poisson$nwspol, c(0.1, 0.9))) # large outliers are removed to indicate variance of male and female

```
The simulated data look similar as the real data based of the survey. The variance is smaller for the simulated one. The higher mean for males are in both plots visible.

### 3.2.4 GLM Quasi-Poisson
```{r quasipoisson}
glmQ_news <- glm(nwspol ~ polintr + eisced + trstprl+ eduyrs + netusoft + stfeco + stfgov +
                   stfdem + gndr + agea + rlgdgr +yrpy,
                 family = "quasipoisson", ## we specify the distribution!
                 data = ess9_poisson)
summary(glmQ_news)
```


```{r quasipoisson - Anova for Gender}

# Checking if Gender plays a role in the quasipoisson model. With the ANOVA function.
glmQ_news_noGender <- update(glmQ_news, . ~ . - gndr)
anova(glmQ_news, glmQ_news_noGender, test = "F")
```
Yes, Gender plays a role for the quasi poisson model. 

```{r quasipoisson - Anova for stfdem}

# Checking if Gender plays a role in the quasipoisson model. With the ANOVA function.
glmQ_news_noGender <- update(glmQ_news, . ~ . - stfdem)
anova(glmQ_news, glmQ_news_noGender, test = "F")
```
To compare it with the variable of "stfdem" about how satisfied the participants of the survey are about how the democracy in their country works. This variable doensn't play a role to predict the media consumption according to the quasipoisson model.


### Labs Example

```{r}

```


<<<<<<< HEAD:ESS9.csv/ESS_DE.Rmd
## 3.3 GLM Binominal
=======
```{r}

```


## 3.3 GLM Binominal 
>>>>>>> d33abb64bd4520b2c184cf8af74b866bfe227af9:ESS9.csv/rfile/ESS_DE.Rmd

## 3.4 GAM

```{r GAM}

ess9_GAM <- ess9_prep

# Multi plots for News Consumption

pairs((nwspol) ~ . , data = ess9_GAM, pch = ".", upper.panel = panel.smooth)
#ploting news consumption and highest level of education (standardized)
ggplot(data = ess9_GAM ,mapping = aes(y = nwspol ,
                                  x = eisced)) +
  geom_boxplot()


# ploting news and age of participants
ggplot(data = ess9_GAM ,mapping = aes(y = nwspol ,
                                  x =  (agea))) +
  geom_point()+
  geom_smooth()

#plotting news and education

ggplot(data = ess9_GAM ,mapping = aes(y = nwspol ,
                                  x =  (eduyrs))) +
  geom_point()+
  geom_smooth()


ggplot(data = ess9_GAM ,mapping = aes(y = nwspol ,
                                  x =  (grspnum))) +
  geom_point()+
  geom_smooth()

ess9_gam <- gam((nwspol) ~ s(eduyrs, k= 1) + s(gndr, k= 1) + s(agea, k= 1) +
                  s(netustm, k= 1) + s(grspnum, k= 1),
                data = ess9_poisson)
data("ozone")
head(ozone)
```


## 3.5 Neural Network 

###3.5.1 continious variable


```{r}

ess9_ann<- ess9_prep_num

str(ess9_ann)

```


#### Prepare for Training

```{r}
set.seed(123)
ess9_ann[factors] <- lapply(ess9_ann[factors], as.numeric)

indices <- createDataPartition(ess9_ann$nwspol, p = 0.8, list = FALSE)
train <- ess9_ann %>% slice(indices)
test <- ess9_ann %>% slice(-indices)
boxplot(train$nwspol, test$nwspol, ess9_ann %>% sample_frac(0.2) %>% pull(nwspol), outline = FALSE)
```

And this time we will also scale the data (by hand for educational purposes)

```{r}
max <- apply(ess9_ann, 2, max)
min <- apply(ess9_ann, 2, min)
ess9_ann_scaled <- as.data.frame(scale(ess9_ann, center = min, scale = max - min))
train_scaled <- ess9_ann_scaled %>% slice(indices)
test_scaled <- ess9_ann_scaled %>% slice(-indices)
```

#### Fit the Network

```{r}


set.seed(42)
ess9_nnet = neuralnet(nwspol ~ polintr + eisced + trstprl+ eduyrs + netusoft + 
                        stfeco + stfgov + stfdem + gndr + agea + rlgdgr +yrpy,
                      train_scaled, hidden = 5 , linear.output = FALSE)
plot(ess9_nnet)
```
#### Predict the test set

```{r, include=FALSE}
pred <- compute(ess9_nnet, test %>% select(-nwspol))
pred$net.result
```
So what went wrong here? Obviously we also need to use the scaled data to make predictions and then scale back to real nwspol values

```{r, include=FALSE}
pred_scaled <- compute(ess9_nnet, test_scaled %>% select(-nwspol))
pred <- pred_scaled$net.result * (max(ess9_ann$nwspol) - min(ess9_ann$nwspol)) + min(ess9_ann$nwspol)
pred
```


```{r}
plot(test$nwspol, pred, col='blue', pch=16, ylab = "predicted nwspol NN", xlab = "real nwspol")
abline(0,1)
```
And calculate the RMSE
```{r}
sqrt(mean((test$nwspol - pred)^2))
```

#### Redo with `caret` and Cross Validation

In principle we cannot be sure that we were not simply "lucky" with the train/test split above, so the proper way to run this would be via `caret` using k-fold Cross Validation. And also as we use `caret` anyway let's also run some hyperparameter tuning:

```{r message=FALSE, warning=FALSE}
set.seed(42)
tuGrid <- expand.grid(.layer1=c(1:4), .layer2=c(0,2), .layer3=c(0))


trCtrl <- trainControl(
  method = 'repeatedcv', 
  number = 4, 
  repeats = 8, 
  returnResamp = 'final'
)

models <- train(
  x = ess9_ann %>% select(-nwspol),
  y = ess9_ann_scaled %>% pull(nwspol),
  method = 'neuralnet', metric = 'RMSE', 
  linear.output = TRUE,
  # be careful, does only work on x!
  preProcess = c('center', 'scale'),
  tuneGrid = tuGrid,
  trControl = trCtrl
)
```


```{r message=FALSE, warning=FALSE}
plot(test$nwspol, pred, col='blue', pch=16, ylab = "predicted rating NN", xlab = "real rating")
abline(0,1)
```

```{r}
plot(models)
```
and extract the best model

```{r}
plot(models$finalModel)
```

###3.5.2 Categorical variable

```{r}
ess9_ann2<- ess9_prep_num

ess9_ann2$netusoft <- as.factor(ess9_ann2$netusoft)


str(ess9_ann2)
```
#### Have a quick look at the Data

```{r}
ess9_ann2 %>%
  ggplot(aes(x = yrpy, y = nwspol, color = netusoft)) +
  geom_point()
```

```{r}
ess9_ann2 %>%
  ggplot(aes(x = eduyrs, y = nwspol, color = netusoft)) +
  geom_point()
```

#### Prepare the Data for Training

```{r}
set.seed(123)
indices <- createDataPartition(ess9_ann2$netusoft, p=.85, list = F)
```

#### Why do we use the `caret` function

Look at the distribution of our different prediction classes in the train and test datasets:

```{r}
ess9_ann2 %>%
  mutate(train = row_number() %in% indices) %>%
  select(netusoft, train) %>%
  table()
```
Now compare this to the "random" approach:

```{r}
ess9_ann2 %>%
  mutate(train = runif(nrow(ess9_ann2)) < .85) %>%
  select(netusoft, train) %>%
  table()
```
So using the `createDataPartition` function makes sure that no class is over- or underrepresented relative to the total occurance in the two sets

### Create some easy Variables to access Data

```{r}
train <- ess9_ann2 %>%
  slice(indices)
test_in <- ess9_ann2 %>%
  slice(-indices) %>%
  select(-netusoft)
test_truth <- ess9_ann2 %>%
  slice(-indices) %>%
  pull(netusoft)
```

## Train the Neural Network

Call the `neuralnet` function creating a network with two hidden layers containing 4 and 3 neurons (probably way too complex for our problem here).

```{r}
set.seed(21)
ess9_ann_net <- neuralnet(netusoft ~ ., train, hidden = c(2,1))
```

Plot the resulting network including the weights

```{r}
plot(ess9_ann_net)
```
## Make Predictions

```{r}
test_results <- neuralnet::compute(ess9_ann_net, test_in)
test_results$net.result
```

Find class (i.e. output neuron) with the highest probability and convert this back into a factor

```{r}
test_pred <- apply(test_results$net.result, 1, which.max)
test_pred <- factor(levels(test_truth)[test_pred], levels = levels(test_truth))
test_pred
```
## Evaluate the Results

```{r}
conf_matrix <- confusionMatrix(test_truth, test_pred)
conf_matrix
```
## Optimize Network Structure

First we need to remodel the data due to some limitations of `caret`

```{r}
xor <- model.matrix(~ netusoft, ess9_ann2)
xor[,1] <- ifelse(xor[,2] %in% 0 & xor[,3] %in% 0, 1, 0)
colnames(xor)[1] <- 'netusoft'
head(xor)
```
```{r}
ess9_ann_mod <- cbind(xor, ess9_ann2 %>% select(-netusoft))
train_mod <- ess9_ann_mod %>% slice(indices)
```


```{r}

set.seed(142)
formula = netusoft + netusoft2 + netusoft3 + netusoft4 + netusoft5 ~ gndr + agea + yrpy + eduyrs
models <- train(formula, train_mod,
               method="neuralnet",   
               ### Parameters for layers
               tuneGrid = expand.grid(.layer1=c(2:4), .layer2=c(0:4), .layer3=c(0)),               
               ### Parameters for optmisation
               learningrate = 0.1,  
               threshold = 0.01,     
               stepmax = 5000         
)
```
And have a look at the different models

```{r}
plot(models)
```


And try out the best model

```{r}
set.seed(42)
best_model <- neuralnet(
  formula,
  ess9_ann_mod %>% slice(indices),
  hidden = c(2,3),
  learningrate = 0.01, 
  threshold = 0.01,   
  stepmax = 50000 
)
plot(best_model)
```

## 3.6 Support Vector Machine
