---
title: "European Survey - What is the political sentiment in German speaking Countries Before Covid?"
author: "Milica Pajkic and Marco Rieder"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

# Getting the Data and installing libraries

```{r, include=FALSE}
if (!require("naniar")) install.packages("naniar", dependencies=TRUE)
if (!require("ggpubr")) install.packages("ggpubr", dependencies=TRUE)
if (!require("cowplot")) install.packages("cowplot", dependencies=TRUE)
library("ggpubr")
library(lubridate)
library(multcomp)
library(dplyr)
library(readr)
library(data.table)
library(tidyverse)
library(naniar)
library(cowplot)
```

```{r}
ess9 <- read_csv("ESS9.csv")
```

```{r making subsets of the whole dataset to only german speaking countries}
ess9_de <- subset(ess9, cntry == "DE", select =dweight:domain)
ess9_ch <- subset(ess9, cntry == "CH", select =dweight:domain)
ess9_at <- subset(ess9, cntry == "AT", select =dweight:domain)
```

# 1. Introduction

# 2. Descriptive Statistics

```{r}
str(ess9_de)
```

## 2.1 Cleaning of the data

```{r Your [pay/pensions/social benefits], which frequency do you know best}
#1 = weekly, 2 = monthly, 3 = annual
ess9_de$infqbst.na <- ess9_de$infqbst
ess9_de["infqbst.na"][ess9_de["infqbst.na"] >= 4] <- NA
ess9_de$infqbst.fac <- factor(ess9_de$infqbst.na)
summary(ess9_de$infqbst.fac)

##Grosspay 
ess9_de$grspnum.na <- ess9_de$grspnum
ess9_de["grspnum.na"][ess9_de["grspnum.na"] >= 666666666] <- NA
summary(ess9_de$grspnum.na)
###Multiplication
ess9_de$yearpay <- ifelse(ess9_de$infqbst %in% 1,
                          ess9_de$grspnum*52,
                          ifelse(ess9_de$infqbst %in% 2,
                                 ess9_de$grspnum*12,
                                 ess9_de$grspnum))
```

```{r recode Variable resp. clean}
#NEWS Politics time recoded
ess9_de$nwspol.na <- ess9_de$nwspol
ess9_de["nwspol.na"][ess9_de["nwspol.na"] >= 7777] <- NA
#-----------------------------------------------------------------------------------------
#Educationlevel recoding
ess9_de$eduyrs.na <- ess9_de$eduyrs
ess9_de["eduyrs.na"][ess9_de["eduyrs.na"] >= 77] <- NA
#Added "Other" also under NA bc we cannot explain what other is
#-----------------------------------------------------------------------------------------
#recode the age variable
ess9_de$age.na <- ess9_de$agea
ess9_de["age.na"][ess9_de["age.na"] >= 140] <- NA
#-----------------------------------------------------------------------------------------
#record the gender variable
ess9_de$gender <- ess9_de$gndr
ess9_de["gender"][ess9_de["gender"] >= 9] <- NA
##factor/categorical
ess9_de$gndr.fac <- factor(ess9_de$gender)
#-----------------------------------------------------------------------------------------
#clean the satisfaction with democracy works in country
ess9_de$stfdem.na <- ess9_de$stfdem
ess9_de["stfdem.na"][ess9_de["stfdem.na"] > 10] <- NA
##factor/categorical
ess9_de$sat.dem.fac <- factor(ess9_de$stfdem.na)
summary(ess9_de$sat.dem.fac)
#-----------------------------------------------------------------------------------------
#clean Boycott certain products
ess9_de$bctprd.na <- ess9_de$bctprd
ess9_de["bctprd.na"][ess9_de["bctprd.na"] > 2] <- NA
##factor/categorical
ess9_de$boycott.fac <- factor(ess9_de$bctprd.na)
#-----------------------------------------------------------------------------------------
#clean left-right scale 0 =l / 10 = r
ess9_de$polscale.na <- ess9_de$lrscale
ess9_de["polscale.na"][ess9_de["polscale.na"] > 10] <- NA
#-----------------------------------------------------------------------------------------
#clean fair chance to participate in politics in yr country (5er scale)
ess9_de$fairchance <- ess9_de$frprtpl
ess9_de["fairchance"][ess9_de["fairchance"] > 6] <- NA
##factor/categorical
ess9_de$fairchance.fac <- factor(ess9_de$fairchance, levels=1:5, labels=c("not at all", "very little", "some", "a lot", "a great deal"))
#-----------------------------------------------------------------------------------------
#clean polinteresse (4er scale)
ess9_de$polinteress <- ess9_de$polintr
ess9_de["polinteress"][ess9_de["polinteress"] > 6] <- NA
##factor/categorical
ess9_de$interest.fac <- factor(ess9_de$polinteress, levels=1:4, 
                               labels=c("very", "quite", "hardly", "none at all"))
###order differently
ess9_de$interest.fac <- ordered(ess9_de$interest.fac, 
                                levels = c("none at all", "hardly", "quite", "very"))
#-----------------------------------------------------------------------------------------
#clean confidence in own ability to participate (5er scale)
ess9_de$ability <- ess9_de$cptppola
ess9_de["ability"][ess9_de["ability"] > 6] <- NA
##factor/categorical
ess9_de$ability.fac <- factor(ess9_de$ability, levels=1:5, labels=c("not at all", "a little", "quite", "very", "completely"))
summary(ess9_de$ability.fac)
#-----------------------------------------------------------------------------------------
#1 = yes / 2 = no / 3 = not eligble
ess9_de$vote.fac <- factor(ess9_de$vote, levels = 1:3, labels = c("Yes", "No", "Not eligible"))
ess9_de["vote.fac"][ess9_de["vote.fac"] >= 4] <- NA
summary(ess9_de$vote.fac)
#-----------------------------------------------------------------------------------------
#post something online in the last 12 month
ess9_de$postonline <- factor(ess9_de$pstplonl)
ess9_de["postonline"][ess9_de["postonline"] > 5] <- NA
ess9_de$postonline<-ifelse(ess9_de$postonline==1,1,0)
summary(ess9_de$postonline)
summary(as.factor(ess9_de$postonline))
summary(as.factor(ess9_de$pstplonl))
#-----------------------------------------------------------------------------------------
##Political system allows people to have a say in what government does 1 = no
ess9_de$gov.allows <- ess9_de$psppsgva
ess9_de["gov.allows"][ess9_de["gov.allows"] > 5] <- NA
ess9_de$gov.allows <- factor(ess9_de$gov.allows)
summary(ess9_de$gov.allows)
#-----------------------------------------------------------------------------------------
#Trust in country's parliament 0 = no trust
ess9_de$trustparl <- ess9_de$trstprl
ess9_de["trustparl"][ess9_de["trustparl"] > 10] <- NA
summary(as.factor(ess9_de$trustparl))
```

```{r}
boxplot(ess9_de$nwspol.na, ess9_de$gndr.fac, outline = FALSE)
```

### wieso ist gender = 2 so komisch?

# 3. Models

## 3.1 Linear Model

Question: What is an influence on the

```{r all variables that i want to include}
#DV
summary(ess9_de$nwspol.na)
#predictorvariables
summary(ess9_de$eduyrs.na)
summary(ess9_de$gndr.fac)
summary(ess9_de$age.na)
summary(ess9_de$stfdem.na)
summary(ess9_de$sat.dem.fac)
summary(ess9_de$ability)
summary(ess9_de$ability.fac)
summary(ess9_de$interest.fac)
summary(as.factor(ess9_de$polinteress))
```

```{r create a DV where no outlier}
ess9_de$nwspol.outlier <- ess9_de$nwspol.na
ess9_de["nwspol.outlier"][ess9_de["nwspol.outlier"] >= 400] <- NA
```

```{r create a subset of data with only the variables mentioned above}
ess9_linear <- ess9_de[, c("nwspol.na", "age.na", "gndr.fac", "eduyrs.na", 
                           "stfdem.na", "ability", "ability.fac", "polinteress", "interest.fac", "sat.dem.fac")]
ess9_linear <- ess9_linear[complete.cases(ess9_linear),]
```

### Graphical analysis

```{r boxplots for the categorial variables w/ outliers}
par(mfrow=c(1,3))
boxplot(nwspol.na ~ gndr.fac, data = ess9_linear)
boxplot(nwspol.na ~ ability.fac, data = ess9_linear)
boxplot(nwspol.na ~ interest.fac, data = ess9_linear)
```

```{r boxplots for the categorial variables w/out outliers}
par(mfrow=c(1, 3))
boxplot(nwspol.na ~ gndr.fac, data = ess9_linear, outline = FALSE)
boxplot(nwspol.na ~ ability.fac, data = ess9_linear, outline = FALSE)
boxplot(nwspol.na ~ interest.fac, data = ess9_linear, outline = FALSE)
```

```{r}
par(mfrow=c(1, 3))
plot()
```

### Fitting Model

```{r linear Regression}
lm_news.0 <- lm(nwspol.na ~ eduyrs.na, data = ess9_linear)
summary(lm_news.0)
```

```{r}
coef(lm_news.0)
```

```{r}
summary(lm_news.0)$coefficients
```

```{r}
par(mfrow=c(1,2))
plot(nwspol.na ~ eduyrs.na, data = ess9_linear,
     xlim = c(0,31), 
     ylim = c(0, 500))
#the ylim was chosen what makes the most sense over 600min watching news equals to 8h/daily
grid(ny = 20)
##
abline(lm_news.0,
       col = "red", lty = "dashed")
#Zoom in 
plot(nwspol.na ~ eduyrs.na, data = ess9_linear,
     xlim = c(0,31), 
     ylim = c(0, 100))
##
grid()
##
abline(lm_news.0,
       col = "red", lty = "dashed")
```

Interpretation: we can see that the line is nearly flat and the p-value confirms this observation. The p-Value is nearly 1 and the parameter is nearly 0.

```{r outlier will be eliminated (above 400min)}
lm.news.outlier <- lm(nwspol.outlier ~ eduyrs.na, data = ess9_de, na.omit = TRUE)
summary(lm.news.outlier)
```

#### Adding additional predictors

```{r}
lm_news.1 <- lm(nwspol.na ~ eduyrs.na + age.na, data = ess9_linear)
summary(lm_news.1)
```

```{r}
coef(lm_news.1)
```

```{r}
summary(lm_news.1)$coefficients
```

```{r}
confint(lm_news.1)
```

We can see that the predictor age has a low p-value and it is significant.

```{r}
lm_news.2 <- lm(nwspol.na ~ eduyrs.na + age.na + gndr.fac + stfdem.na + ability.fac + interest.fac, data = ess9_linear)
summary(lm_news.2)
```

```{r}
coef(lm_news.2)
```

```{r}
summary(lm_news.2)$coefficients
```

```{r}
confint(lm_news.2)
```

```{r}
cis.lm_news.2 <- confint(lm_news.2)
##
## 2) plot estimates
par(mar = c(4,9,2,2))
plot(y = 1:12,
     x = rev(coef(lm_news.2)),
     xlim = c(-20, 80),
     xlab = "Estimated coefficients",
     ylab = "",
     axes = FALSE)
box()
axis(side = 2, at = 1:12,
     labels = rev(names(coef(lm_news.2))), 
     las = 2)
axis(side = 1)

segments(x0 = rev(cis.lm_news.2[, "2.5 %"]),
         x1 = rev(cis.lm_news.2[, "97.5 %"]),
         y0 = 1:12,
         y1 = 1:12)
abline(v = 0, lty = "dashed")
```

#### Interaction: Age \* Interest

```{r}
lm_news.3 <- lm(nwspol.na ~ eduyrs.na + age.na * interest.fac + gndr.fac + stfdem.na + ability.fac, data = ess9_linear)
summary(lm_news.3)
```

```{r}
coef(lm_news.3)
```

```{r}
summary(lm_news.3)$coefficients
```

```{r}
confint(lm_news.3)
```

```{r}
cis.lm_news.3 <- confint(lm_news.3)
##
## 2) plot estimates
par(mar = c(4,10,2,2))
plot(y = 1:15,
     x = rev(coef(lm_news.3)),
     xlim = c(-20, 80),
     xlab = "Estimated coefficients",
     ylab = "",
     axes = FALSE)
box()
axis(side = 2, at = 1:15,
     labels = rev(names(coef(lm_news.3))), 
     las = 2)
axis(side = 1)

segments(x0 = rev(cis.lm_news.3[, "2.5 %"]),
         x1 = rev(cis.lm_news.3[, "97.5 %"]),
         y0 = 1:15,
         y1 = 1:15)
abline(v = 0, lty = "dashed")
```

The interaction is not significant

### Fitted Values and Residuals

#### Fitted Values

```{r fitted values}
fit_lm_news.0 <- fitted(lm_news.0)
str(fit_lm_news.0)
head(fit_lm_news.0)
x.fit.0 <- unique(fit_lm_news.0)
```

```{r fitted values plot lm1}
par(mfrow=c(1,2))
plot(nwspol.na ~ eduyrs.na, data = ess9_linear, 
     main = "Minutes spent on Political News", 
     col = "darkgrey",
     ylim = c(0, 500))
points(fit_lm_news.0 ~ eduyrs.na, 
       col = "purple", 
       pch = 19, 
       data = ess9_linear)
abline(lm_news.0, col = "black")

plot(nwspol.na ~ eduyrs.na, data = ess9_linear, 
     main = "Minutes spent on Political News", 
     col = "darkgrey",
     ylim = c(0, 80))
points(fit_lm_news.0 ~ eduyrs.na, 
       col = "purple", 
       pch = 19, 
       data = ess9_linear)
abline(lm_news.0, col = "black")
```

There are 67 unique values, therefore, the other observations are overlapping. Here we zoomed in to see the negative relationship between news consumption and education years.

#### Residuals

```{r}
resid.news.0 <- resid(lm_news.0)
##
length(resid.news.0 )
head(resid.news.0 )
```

```{r}
set.seed(20) ## for reproducibility 
id <- sample(x = 1:2296, size = 20) 
resid.news.0[id]
```

```{r}
fit_lm_news.0[id]
```

```{r}
plot(nwspol.na ~ eduyrs.na, data = ess9_linear,
     main = "Minutes spent on Political News",
     col = "lightgray",)
##
abline(lm_news.0)
##
points(nwspol.na ~ eduyrs.na, data = ess9_linear[id, ], col = "red")
##
segments(x0 = ess9_linear[id, "eduyrs.na"], x1 = ess9_linear[id,"eduyrs.na"],
         y0 = fit_lm_news.0[id], y1 = ess9_linear[id, "nwspol.na"],
         col = "blue")
```

### Predicted values

```{r}
new.data.news <- data.frame(eduyrs.na = c(18, 7, 28))
pred.new.news <- predict(object = lm_news.0, newdata = new.data.news)

plot(nwspol.na ~ eduyrs.na, data = ess9_linear,
     main = "Minutes spent on Political News",
     col = "lightgray",
      ylim=c(0, 200))
abline(lm_news.0)
points(x = new.data.news$eduyrs.na,
       y = pred.new.news,
       col = "purple",
       pch = 19, cex = 1.5)
```

```{r}
pred.new.news.ci <- predict(object = lm_news.0,
                            interval = "prediction",
        newdata = new.data.news)
pred.new.news.ci
```

```{r}
plot(nwspol.na ~ eduyrs.na, data = ess9_linear,
     main = "Minutes spent on Political News",
     col = "lightgray",
      ylim=c(0, 800))
abline(lm_news.0)
##
points(x = new.data.news$eduyrs.na,
       y = pred.new.news.ci[, "fit"], 
       col = "purple",
       pch = 19, cex = 1.5)
##
segments(x0 = new.data.news$eduyrs.na, 
         x1 = new.data.news$eduyrs.na,
         y0 = pred.new.news.ci[, "lwr"],
         y1 = pred.new.news.ci[, "upr"],
         lwd = 2,
         col = "purple")
```

## 3.2 GLM Poisson

```{r}
glmP_news <- glm(nwspol ~ edu,
family = "poisson", ## we specify the distribution!
data = ess9_ch)
```

## 3.3 GLM Binominal

```{r subset of dataframe because of differen NA}
ess9_binary <- ess9_de[, c("postonline", "age.na", "interest.fac", "gov.allows", "trustparl", "polscale.na", "eduyrs.na", "ability.fac", "sat.dem.fac", "boycott.fac", "vote.fac")]
ess9_binary.na <- ess9_binary[complete.cases(ess9_binary),]
```

```{r}
str(ess9_binary.na)
```

### Graphical Analysis

```{r}
pairs(postonline ~ age.na + interest.fac + gov.allows +
trustparl + polscale.na +
ability.fac + sat.dem.fac + boycott.fac + vote.fac, data = ess9_binary.na, upper.panel = panel.smooth)
```

```{r EDA}

g1 <- ggplot(data = ess9_binary.na,
       mapping = aes(y = postonline,
                     x = age.na))+
  geom_point()+
  geom_smooth(method = "glm", 
              se = FALSE,
              method.args = list(family = "binomial")) 

g2 <- ggplot(data = ess9_binary.na,
       mapping = aes(y = postonline,
                     x = eduyrs.na))+
  geom_point()+
  geom_smooth(method = "glm", 
              se = FALSE,
              method.args = list(family = "binomial")) 

g3 <- ggplot(data = ess9_binary.na,
       mapping = aes(y = postonline,
                     x = interest.fac))+
  geom_point()+
  geom_smooth(method = "glm", 
              se = FALSE,
              method.args = list(family = "binomial")) 

g4 <- ggplot(data = ess9_binary.na,
       mapping = aes(y = postonline,
                     x = ability.fac))+
  geom_point()+
  geom_smooth(method = "glm", 
              se = FALSE,
              method.args = list(family = "binomial")) 

g5 <- ggplot(data = ess9_binary.na,
       mapping = aes(y = postonline,
                     x = boycott.fac))+
  geom_point()+
  geom_smooth(method = "glm", 
              se = FALSE,
              method.args = list(family = "binomial")) 

g6 <- ggplot(data = ess9_binary.na,
       mapping = aes(y = postonline,
                     x = vote.fac))+
  geom_point()+
  geom_smooth(method = "glm", 
              se = FALSE,
              method.args = list(family = "binomial")) 

plot_grid(g1, g2, g3, g4, g5, g6,
          ncol = 2, nrow = 3)
```

### Fitting the binary model (logistic regression

```{r glm model for binary data}
glm_pol.0 <- glm(postonline ~ age.na, data = ess9_binary.na,
                 family = "binomial")
summary(glm_pol.0)
```

```{r}
glm_pol.1 <- glm(postonline ~ age.na + eduyrs.na + interest.fac, data = ess9_binary.na,
                 family = "binomial")
summary(glm_pol.1)
```

### Estimating the performance of a binary model

```{r}
fitted(glm_pol.0) %>% round(digits = 2)
```

```{r}
fitted_glm.pol.0.disc <- ifelse(fitted(glm_pol.0) < 0.3, 
                                yes = 1, no = 0)
head(fitted_glm.pol.0.disc)
summary(as.factor(fitted_glm.pol.0.disc))
```

```{r}
d.obs.fit.glm.pol <- data.frame(obs = ess9_binary.na$postonline,                            
                                fitted = fitted_glm.pol.0.disc) 
table(d.obs.fit.glm.pol)
```

## 3.4 GAM

## 3.5 Neural Network

## 3.6 Support Vector Machine

### Loading the Packages

First make sure we have all the functionality from the `tidyverse` available:

```{r}
library(tidyverse)
theme_set(theme_bw())
```

We are going to use the `e1071` package for this example and also some helper functions from `caret`

```{r}
library(e1071) 
library(caret)
```

```{r}
ess9_de$scale_cat  <- cut(ess9_de$polscale.na,
              breaks=c(-1, 4.5, 7.5, 10),
              labels=c('left', 'center', 'right'))
summary(ess9_de$scale_cat)
summary(as.factor(ess9_de$polscale.na))
ess9_de$scale_cat <- as.factor(ess9_de$scale_cat)
typeof(ess9_de$scale_cat)
summary(ess9_de$scale_cat)
```

```{r}
d.ess9_supportvec <- ess9_de[, c("age.na", "interest.fac", "scale_cat", "eduyrs.na", "sat.dem.fac", "yearpay", "vote.fac")]
d.ess9_supportvec <- d.ess9_supportvec[complete.cases(d.ess9_supportvec),]
str(d.ess9_supportvec)
```

```{r year pay in K}
d.ess9_supportvec$yearpay.k <- d.ess9_supportvec$yearpay/1000
summary(d.ess9_supportvec$yearpay.k)
summary(d.ess9_supportvec$yearpay)
```

### Have a quick look at the Data

```{r}
d.ess9_supportvec %>%
  ggplot(aes(x = age.na, y = yearpay.k, color = scale_cat)) +
  geom_point() 

d.ess9_supportvec %>%
  ggplot(aes(x = age.na, y = yearpay.k, color = vote.fac)) +
  geom_point()
```

```{r}
d.ess9_supportvec %>%
  ggplot(aes(x = age.na, y = eduyrs.na, color = scale_cat)) +
  geom_point()

d.ess9_supportvec %>%
  ggplot(aes(x = age.na, y = eduyrs.na, color = vote.fac)) +
  geom_point()
```

## Prepare the Data for Training

```{r}
set.seed(123)
indices <- createDataPartition(d.ess9_supportvec$scale_cat, p=.85, list = F)
```

### Create some easy Variables to access Data

```{r}
train <- d.ess9_supportvec %>%
  slice(indices)
test_in <- d.ess9_supportvec %>%
  slice(-indices) %>%
  select(-scale_cat)
test_truth <- d.ess9_supportvec %>%
  slice(-indices) %>%
  pull(scale_cat)
```

## Train the Neural Network

Call the `svm` function using the default `cost = 10` parameter.

```{r}
set.seed(123)
ess_de_svm <- svm(scale_cat ~ ., train, kernel = "linear", scale = TRUE, cost = 10)
```

Of course the visualization is a bit difficult as we have some classification regions in four dimensions.

So plot first the resulting classification for the two Petal leaf dimensions

```{r}
plot(ess_de_svm, train, age.na ~ nwspol.na)
```

And for the Sepal leaf dimensions (of course you need to slice the other dimensions at a reasonable point)

```{r}
plot(iris_svm, train, Sepal.Length ~ Sepal.Width, slice = list(Petal.Length = 4.5, Petal.Width = 1.75))
```

## Make Predictions

```{r}
test_pred <- predict(ess_de_svm, test_in)
table(test_pred)
```

## Evaluate the Results

```{r}
conf_matrix <- confusionMatrix(test_pred, test_truth)
conf_matrix
```

So we have a resulting SVM with a prediction accuracy of around 95% (of course we would also need to do some cross validation to ensure we were not simply "lucky" with the train/test split).

## Try again with other parameters

```{r}
set.seed(123)
iris_svm2 <- svm(Species ~ ., train, kernel = "radial", scale = TRUE, cost = 100)
summary(iris_svm2)
```

```{r}
plot(iris_svm2, train, Petal.Length ~ Petal.Width, slice = list(Sepal.Length = 6, Sepal.Width = 3))
```

```{r}
plot(iris_svm2, train, Sepal.Length ~ Sepal.Width, slice = list(Petal.Length = 4.5, Petal.Width = 1.75))
```

```{r}
test_pred <- predict(iris_svm2, test_in)
table(test_pred)
```

```{r}
conf_matrix <- confusionMatrix(test_pred, test_truth)
conf_matrix
```

So we have here the same accuracy but with much more complex boundaries (radial kernels and higher costs) so we might be overfitting here a bit more.
