---
title: "European Survey - What is the political sentiment in German speaking Countries Before Covid?"
author: "Milica Pajkic and Marco Rieder"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

# Getting the Data and installing libraries

```{r, include=FALSE}
library(lubridate)
library(multcomp)
library(dplyr)
library(readr)
library(data.table)
library(tidyverse)
library(naniar)
library(neuralnet)
library(nnet)
library(caret)
library(gamlss.add)
library(ggplot2)
```

```{r}

ess9 <- read_csv("ESS9.csv")

unique(ess9$name)
unique(ess9$gndr)
unique(ess9$cntry)
```

```{r making subsets of the whole dataset to only german speaking countries}
ess9_de <- subset(ess9, cntry == "DE", select =dweight:domain)
ess9_ch <- subset(ess9, cntry == "CH", select =dweight:domain)
ess9_at <- subset(ess9, cntry == "AT", select =dweight:domain)
```

# 1. Introduction

The European Social Survey (ESS) is a large scale survey conducted in over 38 countries within Europe. Focusing on public attitudes and values and changes within time. This paper is evaluating the ninth version of the survey (ESS9) form 2018.


# 2. Descriptive Statistics

```{r }
summary(ess9_poisson$nwspol)
sum(is.na(ess9_poisson$nwspol))
```

```{r echo=FALSE}
summary(ess9_de$gndr)
sum(is.na(ess9_de$gndr))
typeof(ess9_de$nwspol)
```

```{r}
ggplot(ess9_de, aes(nwspol, 1)) +
  geom_bar(stat = "identity") + 
  xlim(c(0, 200))
```

```{r}
boxplot(ess9_de$nwspol, outline=FALSE)
```

### **Was muss man alles für eine deskriptive Analyse machen?**

```{r lineare Regression}
typeof(ess9_de$eduade3)
summary(ess9_de$eduade3)
sum(is.na(ess9_de$eduade3))
```

# 3. Models

## 3.1 Linear Model

How much politics are you watching = DV\
education level by years = IV\
happy with government = IV\
Gender = IV\
Age = IV


```{r linear Regression}
lm_news.1 <- lm(nwspol ~ eduade3, data = ess9_de)
summary(lm_news.1)
```

## 3.2 GLM Poisson

The media consumption in minutes is count data, therefore the GLM Poisson model can be used and indicate how the 

```{r}
# Setting NA values correct for the used parameters replace with NA replace function from naniar package

ess9_poisson <- ess9 %>%
  replace_with_na(replace = list(nwspol =
                                   c(6666, 7777, 8888, 9999),
                                 polintr = c(7,8,9),
                                 eisced = c(77,88,99),
                                 eduyrs = c(77, 88, 99),
                                 stfeco = c(77, 88, 99),
                                 stfgov = c(77, 88, 99),
                                 stfdem = c(77, 88, 99),
                                 gndr = 9,
                                 agea = 999,
                                 rlgdgr = c(77, 88, 99),
                                 netustm = c(6666, 7777, 8888, 9999),
                                 ggchld = c(6,7,8,9)
                                 ))

#testing if change has worked and how many NA we have - for four examples

sum(is.na(ess9_poisson$nwspol))
sum(is.na(ess9_poisson$eduyrs))
sum(is.na(ess9_poisson$stfgov))
sum(is.na(ess9_poisson$agea))



# drop NA for the columns used for the poisson model
ess9_poisson <- ess9_poisson %>% drop_na((c(
  nwspol,eduyrs, stfeco, stfgov, stfdem, gndr, agea, rlgdgr, polintr)))



ggplot(data = ess9_poisson,
mapping = aes(y = nwspol,
x = eduyrs)) +
geom_hline(yintercept = 0) + ## to highlight the lower bound of the data
geom_boxplot()



# fit the model and use 5 parameter
glmP_news <- glm(nwspol ~ polintr + eduyrs + netusoft + stfeco + stfgov +
                   stfdem + gndr + agea + rlgdgr,
                 family = "poisson", ## we specify the distribution!
                 data = ess9_poisson)

summary(glmP_news)
```

Now we simulate with our model the average media consumption for Females and Males.
```{r}

#simulate the news consumption in minutes value from the model for the two gender

set.seed(2) # for reproducibility
#simulate the data from glm poisson
sim.data.ess9.Poisson <- simulate(glmP_news)
##

NROW(sim.data.ess9.Poisson)
head(sim.data.ess9.Poisson)

#plot data as bar plot, with gender as factor to get two groups to compare, without factoring only one plot would show.
ggplot(mapping = aes(y = sim.data.ess9.Poisson$sim_1,
                     x = as.factor(ess9_poisson$gndr))) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  ylab("simulated media consumption \n(assuming Poisson dist)") +
  xlab("Gender")
```
### 3.2.1 GLM Quasi-Poisson


```{r}

glmQ_news <- glm(nwspol ~ eduyrs + stfeco + stfgov + stfdem + gndr + agea + rlgdgr,
                 family = "quasipoisson", ## we specify the distribution!
                 data = ess9)
summary(glmQ_news)



```


<<<<<<< HEAD:ESS9.csv/ESS_DE.Rmd
## 3.3 GLM Binominal
=======
```{r}

```


## 3.3 GLM Binominal 
>>>>>>> d33abb64bd4520b2c184cf8af74b866bfe227af9:ESS9.csv/rfile/ESS_DE.Rmd

## 3.4 GAM

```{r}

#ploting news consumption and highest level of education (standardized)
ggplot(data = ess9_poisson ,mapping = aes(y = nwspol ,
                                  x = as.factor (eisced))) +
  geom_boxplot()


ggplot(data = ess9_poisson ,mapping = aes(y = nwspol ,
                                  x = as.factor (ggchld))) +
  geom_boxplot()



ggplot(data = ess9_poisson ,mapping = aes(y = nwspol ,
                                  x =  (agea))) +
  geom_point()+
  geom_smooth()

```


## 3.5 Neural Network

```{r message=FALSE, warning=FALSE}
max <- apply(ess9, 2, max)
min <- apply(ess9, 2, min)
```


```{r message=FALSE, warning=FALSE}
#ess9_scaled <- as.data.frame(scale(ess9, center = min, scale = max - min))



set.seed(42)
tuGrid <- expand.grid(.layer1=c(1:4), .layer2=c(0,2), .layer3=c(0))

trCtrl <- trainControl(
  method = 'repeatedcv', 
  number = 5, 
  repeats = 10, 
  returnResamp = 'final'
)

models <- train(
  x = ess9 %>% dplyr::select(-nwspol),
  y = ess9 %>% pull(nwspol),
  method = 'neuralnet', metric = 'RMSE', 
  linear.output = TRUE,
  # be careful, does only work on x!
  preProcess = c('center', 'scale'),
  tuneGrid = tuGrid,
  trControl = trCtrl
)
```

```{r}
plot(models)
```
and extract the best model

```{r}
plot(models$finalModel)
```


## 3.6 Support Vector Machine
