---
title: "European Survey - What is the political sentiment in German speaking Countries Before Covid?"
author: "Milica Pajkic and Marco Rieder"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

# Getting the Data and installing libraries

```{r, include=FALSE}
library(lubridate)
library(multcomp)
library(dplyr)
library(readr)
library(data.table)
library(tidyverse)
library(naniar)
library(neuralnet)
library(nnet)
library(caret)
library(gamlss.add)
library(ggplot2)
library(mgcv)

```

```{r}
ess9 <- read_csv("ESS9.csv")

unique(ess9$name)
unique(ess9$gndr)
unique(ess9$cntry)
```

```{r making subsets of the whole dataset to only german speaking countries}
ess9_de <- subset(ess9, cntry == "DE", select =dweight:domain)
ess9_ch <- subset(ess9, cntry == "CH", select =dweight:domain)
ess9_at <- subset(ess9, cntry == "AT", select =dweight:domain)
```

# 1. Introduction

The European Social Survey (ESS) is a large scale survey conducted in over 38 countries within Europe. Focusing on public attitudes and values and changes within time. This paper is evaluating the ninth version of the survey (ESS9) form 2018.

## 1.1 Selection of parameters

The survey is really comprehensive and consists out of 572 variables. Some of them are related to others and only answered by a subset of participants. 

## 1.2 Goal

# 2. Descriptive Statistics

First 

```{r }
summary(ess9$nwspol)
sum(is.na(ess9$nwspol))
```

```{r echo=FALSE}
summary(ess9_de$gndr)
sum(is.na(ess9_de$gndr))
typeof(ess9_de$nwspol)
```

```{r}
ggplot(ess9_de, aes(nwspol, 1)) +
  geom_bar(stat = "identity") + 
  xlim(c(0, 200))
```

```{r}
boxplot(ess9_de$nwspol, outline=FALSE)
```

### **Was muss man alles für eine deskriptive Analyse machen?**

```{r lineare Regression}
typeof(ess9_de$eduade3)
summary(ess9_de$eduade3)
sum(is.na(ess9_de$eduade3))
```

# 3. Models

## 3.1 Linear Model

How much politics are you watching = DV\
education level by years = IV\
happy with government = IV\
Gender = IV\
Age = IV


```{r linear Regression}
lm_news.1 <- lm(nwspol ~ eduade3, data = ess9_de)
summary(lm_news.1)
```

## 3.2 GLM Poisson

The media consumption in minutes is count data, therefore the GLM Poisson model can be used and indicate how the 

```{r}
# Setting NA values correct for the used parameters replace with NA replace function from naniar package

ess9_poisson <- ess9 %>%
  replace_with_na(replace = list(nwspol =
                                   c(6666, 7777, 8888, 9999),
                                 polintr = c(7,8,9),
                                 eisced = c(77,88,99),
                                 eduyrs = c(77, 88, 99),
                                 stfeco = c(77, 88, 99),
                                 stfgov = c(77, 88, 99),
                                 stfdem = c(77, 88, 99),
                                 gndr = 9,
                                 agea = 999,
                                 rlgdgr = c(77, 88, 99),
                                 netustm = c(6666, 7777, 8888, 9999),
                                 netusoft = c(7,8,9),
                                 ggchld = c(6,7,8,9),
                                 psppsgva = c(7,8,9),
                                 grspnum = c(666666666,777777777,888888888,999999999)


                                 ))

#testing if change has worked and how many NA we have - for four examples

sum(is.na(ess9_poisson$nwspol))
sum(is.na(ess9_poisson$eduyrs))
sum(is.na(ess9_poisson$stfgov))
sum(is.na(ess9_poisson$agea))

# defining subset for study and models
ess9_factor <- c("nwspol", "polintr", "eisced", "eduyrs","stfeco", "stfgov", "stfdem", "gndr", "agea","rlgdgr", "netustm", "netusoft", "ggchld", "psppsgva", "grspnum")

#subsetting the dataset with the defined parameters
ess9_poisson <- ess9_poisson[ess9_factor]

# drop NA for the columns used for the poisson model
ess9_poisson <- ess9_poisson %>% drop_na((ess9_factor))



ggplot(data = ess9_poisson, mapping = aes(y = nwspol,x = as.factor(gndr))) +
  geom_hline(yintercept = 0) + ## to highlight the lower bound of the data
  geom_boxplot()



# fit the model and use 5 parameter
glmP_news <- glm(nwspol ~ polintr + eduyrs + netusoft + stfeco + stfgov +
                   stfdem + gndr + agea + rlgdgr,
                 family = "poisson", ## we specify the distribution!
                 data = ess9_poisson)

summary(glmP_news)
```

Now we simulate with our model the average media consumption for Females and Males.
```{r poisson}

#simulate the news consumption in minutes value from the model for the two gender

set.seed(2) # for reproducibility
#simulate the data from glm poisson
sim.data.ess9.Poisson <- simulate(glmP_news)
##

NROW(sim.data.ess9.Poisson)
head(sim.data.ess9.Poisson)

#plot data as bar plot, with gender as factor to get two groups to compare, without factoring only one plot would show.
ggplot(mapping = aes(y = sim.data.ess9.Poisson$sim_1,
                     x = as.factor(ess9_poisson$gndr))) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  ylab("simulated media consumption \n(assuming Poisson dist)") +
  xlab("Gender")
```
### 3.2.1 GLM Quasi-Poisson


```{r quasipoisson}

glmQ_news <- glm(nwspol ~ eduyrs + stfeco + stfgov + stfdem + gndr + agea + rlgdgr,
                 family = "quasipoisson", ## we specify the distribution!
                 data = ess9)
summary(glmQ_news)



```


<<<<<<< HEAD:ESS9.csv/ESS_DE.Rmd
## 3.3 GLM Binominal
=======
```{r}

```


## 3.3 GLM Binominal 
>>>>>>> d33abb64bd4520b2c184cf8af74b866bfe227af9:ESS9.csv/rfile/ESS_DE.Rmd

## 3.4 GAM

```{r GAM}

# Multi plots for News Consumption

pairs((nwspol) ~ . , data = ess9_poisson, pch = ".", upper.panel = panel.smooth)


#ploting news consumption and highest level of education (standardized)
ggplot(data = ess9_poisson ,mapping = aes(y = nwspol ,
                                  x = as.factor (eisced))) +
  geom_boxplot()

# ploting news and if there are grand childern 
ggplot(data = ess9_poisson ,mapping = aes(y = nwspol ,
                                  x = as.factor (ggchld))) +
  geom_boxplot()


# ploting news and age of participants
ggplot(data = ess9_poisson ,mapping = aes(y = nwspol ,
                                  x =  (agea))) +
  geom_point()+
  geom_smooth()

#plotting news and education

ggplot(data = ess9_poisson ,mapping = aes(y = nwspol ,
                                  x =  (eduyrs))) +
  geom_point()+
  geom_smooth()


ggplot(data = ess9_poisson ,mapping = aes(y = nwspol ,
                                  x =  (grspnum))) +
  geom_point()+
  geom_smooth()

# ess9_gam <- gam((nwspol) ~ s(eduyrs, k= 1) + s(gndr, k= 1) + s(agea, k= 1) +
#                   s(netustm, k= 1) + s(grspnum, k= 1),
#                 data = ess9_poisson)


```


## 3.5 Neural Network


```{r}
str(ess9_poisson)
```


### Prepare for Training

```{r}
set.seed(123)
indices <- createDataPartition(ess9_poisson$nwspol, p = 0.8, list = FALSE)
train <- ess9_poisson %>% slice(indices)
test <- ess9_poisson %>% slice(-indices)
boxplot(train$nwspol, test$nwspol, ess9_poisson %>% sample_frac(0.2) %>% pull(nwspol))
```

And this time we will also scale the data (by hand for educational purposes)

```{r}
max <- apply(ess9_poisson, 2, max)
min <- apply(ess9_poisson, 2, min)
ess9_poisson_scaled <- as.data.frame(scale(ess9_poisson, center = min, scale = max - min))
train_scaled <- ess9_poisson_scaled %>% slice(indices)
test_scaled <- ess9_poisson_scaled %>% slice(-indices)
```

### Fit the Network

```{r}
set.seed(42)
cereal_net = neuralnet(nwspol ~ polintr + eisced + eduyrs + stfeco + stfdem + gndr + agea + rlgdgr + netusoft + ggchld + psppsgva + grspnum, train_scaled, hidden = 3 , linear.output = TRUE)
plot(cereal_net)
```
### Predict the test set

```{r}
pred <- compute(cereal_net, test %>% select(-nwspol))
pred$net.result
```
So what went wrong here? Obviously we also need to use the scaled data to make predictions and then scale back to real nwspol values

```{r}
pred_scaled <- compute(cereal_net, test_scaled %>% select(-nwspol))
pred <- pred_scaled$net.result * (max(ess9_poisson$nwspol) - min(ess9_poisson$nwspol)) + min(ess9_poisson$nwspol)
pred
```


```{r}
plot(test$nwspol, pred, col='blue', pch=16, ylab = "predicted nwspol NN", xlab = "real nwspol")
abline(0,1)
```
And calculate the RMSE
```{r}
sqrt(mean((test$nwspol - pred)^2))
```

### Redo with `caret` and Cross Validation

In principle we cannot be sure that we were not simply "lucky" with the train/test split above, so the proper way to run this would be via `caret` using k-fold Cross Validation. And also as we use `caret` anyway let's also run some hyperparameter tuning:

```{r message=FALSE, warning=FALSE}
set.seed(42)
tuGrid <- expand.grid(.layer1=c(1:4), .layer2=c(0,2), .layer3=c(0))

trCtrl <- trainControl(
  method = 'repeatedcv', 
  number = 5, 
  repeats = 10, 
  returnResamp = 'final'
)

models <- train(
  x = ess9_poisson %>% select(-nwspol),
  y = ess9_poisson_scaled %>% pull(nwspol),
  method = 'neuralnet', metric = 'RMSE', 
  linear.output = TRUE,
  # be careful, does only work on x!
  preProcess = c('center', 'scale'),
  tuneGrid = tuGrid,
  trControl = trCtrl
)
```

```{r}
plot(models)
```
and extract the best model

```{r}
plot(models$finalModel)
```

```{r message=FALSE, warning=FALSE}
#ess9_scaled <- as.data.frame(scale(ess9, center = min, scale = max - min))

set.seed(42)
tuGrid <- expand.grid(.layer1=c(1:4), .layer2=c(0,2), .layer3=c(0))

trCtrl <- trainControl(
  method = 'repeatedcv', 
  number = 5, 
  repeats = 10, 
  returnResamp = 'final'
)

models <- train(
  x = ess9_poisson %>% select(-nwspol),
  y = ess9_poisson_scaled %>% pull(nwspol),
  method = 'neuralnet', metric = 'RMSE', 
  linear.output = TRUE,
  # be careful, does only work on x!
  preProcess = c('center', 'scale'),
  tuneGrid = tuGrid,
  trControl = trCtrl
)
```

```{r}
plot(models)
```
and extract the best model

```{r}
plot(models$finalModel)
```

```{r}

ess9_NN <- nnet(nwspol ~ ., data = ess9_poisson, size=15, maxit=100, range=0.1, decay=5e-4)




```

## 3.6 Support Vector Machine
